climatrends: Precipitation and temperature indices for climate variability analysis in R

### 2020-05-07 inst/paper/paper.Rmd
adjust-travis-yml
fix-methods-in-ETo-and-GDD

Understand how climate variability drives ecological processes is a key approach to provide recommendations for climate adaptation and biodiversity conservation. **climatrends** aims to provide the R [@RCoreTeam] toolkit to compute extreme precipitation and temperature indices that serves as input for climate and crop models [@vanEtten2019; @Kehel2016], trends in climate change [@Aguilar2005; @deSousa2018] and applied ecology [**add-reference**].

# Implementation

@@

# Application: common beans

@@

# Application: time series

@@

# Further development
Integration with other datasets as they become available in `R` via API client packages. New indices related to the physiology of crops to be implemented while I work on the rice data.

>>>>

Understand how climate variability drives ecological processes is a key approach to provide recommendations for climate adaptation and biodiversity conservation. **climatrends** aims to provide the R [@RCoreTeam] toolkit to compute extreme precipitation and temperature indices that serves as input for climate and crop models [@vanEtten2019; @Kehel2016], trends in climate change [@Aguilar2005; @deSousa2018] and applied ecology [@Prentice1992; @YLiu2018].

[...continue...]

# Methods and features

## Implementation

@@

[...continue...]

## Growing degree-days

Growing degree-days (gdd) is an heuristic tool in phenology that measures heat accumulation and is used to predict plant and animal development rates [@Prentice1992]. Growing degree-days are calculated by taking the integral of warmth above a base temperature ($T_{0}$). The function `GDD()` applies by default the following equation.

Equation [1]

$$GDD = \frac{T_{max} + T_{min}}{2} - T_{0}$$

Where $T_{max}$ is the maximum temperature in the given day, $T_{min}$ is the minimum temperature in the given day and $T_{0}$ is the minimum temperature for growth (as per the physiology of the focal organism).

For cold environments, where $T_{min}$ may be lower than $T_{0}$, there are two variants of the previous equation to adjust either $T_{mean}$ (variant A) or $T_{min}$ (variant B). The variant A changes $T_{mean}$ to $T_{0}$ if $T_{mean} < T_{0}$ and is expressed as follows.

Equation [2]

$$ GDD = max \left(\frac{T_{max} + T_{min}}{2} - T_{0}, \; 0 \right)$$

The variant B, is calculated using Equation 1, but adjusts $T_{min}$ to $T_{0}$ if $T_{min} < T_{o}$, and is adjusted as follows.

Equation [3]

$$ T_{min} < T_{o} \; \rightarrow \; T_{min} = T_{o} $$

In `GDD()` these equations are defined as 'default', 'variant_a' and 'variant_b', respectively, and can be selected using the argument `equation`.

Additionally, the function may return the number of days that a given organism required to reach a certain number of accumulated degree-days or the raw gdd in each observed day. These values are defined by 'ndays' or 'gdd' and can be adjusted using the argument `return.as`. The required accumulated gdd is defined with argument `degree.days`. For example, the Korean pine (*Pinus koraiensis*) requires 105 $^\circ C$ accumulated gdd to onset of photosynthesis [@JWu2013].

## Reference evapotranspiration

The reference evapotranspiration measures the influence of the climate on a given organism (generally a crop species) water needs [@Brouwer1986]. The function `ETo()` applies the Blaney-Criddle formula, a general theoretical method used when no measured data on pan evaporation is available locally, using the following equation.

Equation [4]

$$ETo = p \times \left(0.46 \times \frac{T_{max} + T_{min}}{2} + 8 \right) \times K_c$$

Where $p$ is the mean daily percentage of annual daytime hours, $T_{max}$ is the maximum temperature, $T_{min}$ is the minimum temperature, and $K_c$ is the relationship between the reference organism and the actually grown of the organism and may change according to the growth stage.

# Examples

## Common beans

@@

## Time series

@@

# Further development

Integration with other datasets as they become available in `R` via API client packages. New indices related to the physiology of crops to be implemented while I work on the rice data.
